CAD and Routing Architecture for Interposer-Based Multi-FPGA Systems

### 2014-11-30 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
updated figure caption
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4510 8e3573b8-cf2c-4f14-ef6d-137439e28b8b
updated based on Vaughn's comments on Draft 2
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4515 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

%   - routing-interposer interface
\subsection{Routing-Interposer Interface}
As described in \ref{sec:interposer-routing-interface}, we have added three boolean options to VPR that allow us to experiment with different routing-interposer interfaces. These interfaces provide more flexibility to route signals between different dice and hence help us achieve lower \textit{minimum channel width} and \textit{critical path delay}.

To measure the impact of routing-interposer interface parameters, we set \textit{delay increase}=$1ns$, \textit{number of cuts=1}, and \textit{\% wires cut=80\%}; then, we measure the impact of the \textit{bidirectional interposer wires}, \textit{fanin transfer}, and \textit{fanout transfer} parameters in different combinations. Figure~\ref{fig:all_arch_experiments} summarizes the results of these experiments; all results are normalized to the least flexible routing-interposer interface.

\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{arch_experiments2.eps}
\caption{Impact of routing-interposer interface on minimum channel width and circuit speed.}
\label{fig:all_arch_experiments}
\end{figure}

@@

Moreover, better performance can be achieved if either of these two options is used together with bidirectional interposer wires. Enabling \textit{fanout transfer} with bidirectional interposer wires (column \textit{d} in Figure~\ref{fig:all_arch_experiments}) reduces the minimum channel width by 9\% and critical path delay by 0.7\%. Enabling \textit{fanin transfer} with bidirectional interposer wires (column \textit{e} in Figure~\ref{fig:all_arch_experiments}) has an even stronger impact and reduces the minimum channel width by 10\% and critical path delay by 0.7\%. This signifies that the flexibility to drive an interposer wire is more important for routability compared to the flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.

Enabling all of these architecture modifications simultaneously (column \textit{f} in Figure~\ref{fig:all_arch_experiments}) provides the highest level of flexibility for the router and achieves 11.2\% reduction in minimum channel width and 1.3\% reduction in critical path delay. These models show that interposer-based FPGAs can extensively benefit from hardware that provides a flexible routing-interposer interface. It is important both to make the interposer wires easier to access with additional routing switches \textit{and} to make them bidirectional to accommodate asymmetric routing demand.

@@

Figure~\ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}. It can be noted from this graph that the minimum channel width is essentially constant up to 60\% of wires cut, indicating that the CAD flow is able to avoid saturating the interposer routing until that point. When more than 80\% of the wires are cut however (i.e. the interposer provides less than 20\% of the usual within-die routing capacity), the minimum channel width grows rapidly indicating that the interposer routing bandwidth has become a limiting factor.

@@

The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut. Note, however, that this same critical path delay is achieved at a much higher channel width.

>>>>

%   - routing-interposer interface
\subsection{Routing-Interposer Interface}
As described in \ref{sec:interposer-routing-interface}, we have added three boolean options to VPR that allow us to experiment with different routing-interposer interfaces. Some interface choices provide more flexibility to route signals between different dice and hence help us achieve lower \textit{minimum channel width} and \textit{critical path delay}.

To measure the impact of routing-interposer interface parameters, we set \textit{delay increase}=$1ns$, \textit{number of cuts=1}, and \textit{\% wires cut=80\%}; then, we measure the impact of the \textit{bidirectional interposer wires}, \textit{fanin transfer}, and \textit{fanout transfer} parameters in different combinations. Figure~\ref{fig:all_arch_experiments} summarizes the results of these experiments; all results are normalized to the least flexible routing-interposer interface: one in which 80\% of wires end at the interposer and the other 20\% cross the interposer with the same directionality and routing switches they would have in a conventional FPGA.

\begin{figure}[!h]
\centering
\includegraphics[width=\linewidth]{arch_experiments3.eps}
\caption{Impact of routing-interposer interface flexibility on minimum channel width and circuit speed.}
\label{fig:all_arch_experiments}
\end{figure}

@@

Moreover, better performance can be achieved if either of these two options is used together with bidirectional interposer wires. Enabling \textit{fanout transfer} with bidirectional interposer wires (column \textit{d} in Figure~\ref{fig:all_arch_experiments}) reduces the minimum channel width by 9\% and critical path delay by 0.7\%. Enabling \textit{fanin transfer} with bidirectional interposer wires (column \textit{e} in Figure~\ref{fig:all_arch_experiments}) has an even stronger impact and reduces the minimum channel width by 10\% and critical path delay by 0.7\%. This signifies that added flexibility to drive an interposer wire is more important for routability compared to adding flexibility of having an interposer wire drive multiple vertical routing wires on the other side of the cut.

Enabling all of these architecture modifications simultaneously (column \textit{f} in Figure~\ref{fig:all_arch_experiments}) provides the highest level of flexibility for the router and achieves 11.2\% reduction in minimum channel width and 1.3\% reduction in critical path delay. These models show that interposer-based FPGAs significantly benefit from hardware that provides a flexible routing-interposer interface. It is important both to make the interposer wires easier to access with additional routing switches \textit{and} to make them bidirectional to accommodate asymmetric routing demand.

@@

Figure~\ref{fig:standard_minW} shows the graph of minimum channel width versus \textit{\% wires cut}. It can be noted from this graph that the minimum channel width is essentially constant up to 60\% of wires cut, indicating that the CAD flow is able to avoid saturating the interposer routing until that point. When more than 80\% of the wires are cut however (i.e. the interposer provides less than 20\% of the usual within-die routing capacity), the minimum channel width grows rapidly, indicating that the interposer routing bandwidth has become a limiting factor.

@@

The critical path delay, depicted in Figure~\ref{fig:standard_crit}, on the other hand, is not strongly influenced by the percentage of wires cut, as the critical path delay at 80\% of the wires cut is essentially the same as at 0\% wires cut. Note, however, that this same critical path delay is achieved at a much higher channel width when \textit{\% wires cut} is greater than 80\%.
